---
title: "class15"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Hands on section
##Import countData and colData into R
```{r}
counts <- read.csv("data/airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <- read.csv("data/airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
dim(counts)
head(counts)
```
```{r}
head(metadata)
```
```{r}
colnames(counts)
```
```{r}
metadata$id
```

Check for corresponance of metadata and countdata
```{r}
colnames(counts)[-1] ==metadata$id
#then check to see if they are all true
all(colnames(counts)[-1] == metadata$id)
```


First task find which columns in **counts** correspond to the control conditions (i.e. no drug!). Then we can calculate the mean value for each gene.
```{r}
control.inds <- metadata$dex == "control"
control.ids <- metadata$id[control.inds]

head(counts[,control.ids])

```

```{r}
rowSums(counts[,control.ids]) / ncol(counts[,control.ids])
```


Now we are doing the same as above but with treated instead of control

```{r}
treated.inds <- metadata$dex == "treated"
treated.ids <- metadata$id[treated.inds]

head(counts[,treated.ids])
```
```{r}
rowSums(counts[,treated.ids]) / ncol(counts[,treated.ids])
```

```{r}
control.mean <- rowSums(counts[,control.ids]) / ncol(counts[,control.ids])
names(control.mean) <- counts$ensegene
head(control.mean)
```

```{r}
control <- metadata[metadata[,"dex"]=="control",]
control.mean <- rowSums( counts[ ,control$id] )/4 
names(control.mean) <- counts$ensgene
```

Now, same thing with treated colummns
```{r}
treated <- metadata[metadata[,"dex"]=="treated",]
treated.mean <- rowSums( counts[ ,treated$id] )/ncol(counts[,treated$id])
names(control.mean) <- counts$ensgene
```

Combine the mean count values for control and treated into a new data.frame called meancounts!
```{r}
meancounts <- data.frame(control.mean, treated.mean)
```

```{r}
colSums(meancounts)
```

Now we can plot the drug treater vs non drug...
```{r}
plot(meancounts$control.mean, meancounts$treated.mean, log = "xy")
```

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
#NaN means answer is not a number - log of 
```


```{r}
toy <- c(1, 2, 0, 4, 0)
toy==0
which(toy==0)
```

```{r}
toy <- meancounts[1:6,]
toy[,1:2] == 0
which(toy[,1:2]==0, arr.ind = TRUE)

which(toy[,1:2]==0, arr.ind = TRUE)[,1]
```

```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```

```{r}
sum(up.ind)
sum(down.ind)
```

There are `sum(up.ind)` up regulated genes and `sum(down.ind)` down regulated genes. 

```{r}
head(mycounts[up.ind,])
```

##Adding annotation data

Use the **merge()** function to add annotation data from ensemble

```{r}
anno <- read.csv("data/annotables_grch38.csv")
head(anno)
```

```{r}
mycounts.annotated <- merge(x=mycounts, y=anno, by.x="row.names", by.y="ensgene")
head(mycounts.annotated)
```


##DESeq2 analysis

Now we do this analysis properly with p-values!
```{r}
library(DESeq2)
citation("DESeq2")
```

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
head(res)
```

```{r}
summary(res)
```

#Making a plot (Volcano plot)
```{r}
res$sig <- res$padj<0.05

# How many of each?
table(res$sig)
```
```{r}
sum(is.na(res$sig))
```

```{r}
#Writing code for plot
mycols <- rep("gray", nrow(res))
mycols[abs(res$log2FoldChange)>2] <- "red"

inds <- (res$padj<0.01) & (abs(res$log2FoldChange)>2)
mycols[inds] <- "blue"

plot(res$log2FoldChange, -log(res$padj), col=mycols)
abline(h=-log(0,1), col="gray", lty=2)
```


```{r}
#Code from class worksheet
library(ggplot2)

ggplot(as.data.frame(res), aes(log2FoldChange, -1*log10(pvalue), col=sig)) + 
    geom_point() + 
    ggtitle("Volcano plot")
```

