---
title: "class09"
author: "Reshma Pyala"
date: "2/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
 ##Preparing the data
 
 Adding data to the name fna.data
```{r}
# Save your input data file to a new 'data' directory
fna.data <- "WisconsinCancer.csv"

# Complete the following code to input the data and store as wisc.df
wisc.df <- read.csv(fna.data)
head(wisc.df)
```

```{r}
View(wisc.df)

```

```{r}
nrow(wisc.df)
```

```{r}
colnames(wisc.df)
#Doing this to see the column names and the number of column that deals with the "X" that is unwanted.
```

Next use as.matrix() to convert the other features (i.e. columns) of the data (in columns 3 through 32) to a matrix. Store this in a variable called wisc.data.
```{r}
# Convert the features of the data: wisc.data
wisc.data <- as.matrix(wisc.df[,3:32])
```

Assign the row names of wisc.data the values currently contained in the id column of wisc.df. While not strictly required, this will help you keep track of the different observations throughout the modeling process.
```{r}
# Set the row names of wisc.data
row.names(wisc.data) <- wisc.df$id
#Set as dollar id bc that was initially the column with the names of each so we just adjusted the data to be better fit for formatting
#head(wisc.data)
View(wisc.data)
```

How many cancer (M) and non cancer samples so we have in our data?
```{r}
table(wisc.df$diagnosis)

```

Finally, setup a separate new vector called diagnosis to be 1 if a diagnosis is malignant ("M") and 0 otherwise. Note that R coerces TRUE to 1 and FALSE to 0.
```{r}
# Create diagnosis vector by completing the missing code
#diagnosis <- as.numeric(wisc.df$diagnosis)
diagnosis <- as.numeric(wisc.df$diagnosis == "M")
```

Q1. How many observations are in this dataset?
```{r}
nrow(wisc.data)
```

Q2. How many variables/features in the data are suffixed with _mean?
```{r}
#useful to keep:
  #grep("_pattern", x)
  #prcomp(x, scale = TRUE)
grep("_mean", colnames(wisc.data))
inds <- grep("_mean", colnames(wisc.data))
colnames(wisc.data[,1:10])

length(inds)
```

Q3. How many of the observations have a malignant diagnosis?
```{r}
sum(diagnosis)
```


##Principal Component Analysis

```{r}
# Check column means and standard deviations
colMeans(wisc.data)

apply(wisc.data,2,sd)
```

Let's do PCA
```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data, scale = TRUE)
summary(wisc.pr)
```

Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?
  0.4427 from PC1

Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
  Would need 3 components to meet 70% mark. Would need 5 components to meet 80% mark.

Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?
  Would need 7 components to meet 90% mark.
  
```{r}
biplot(wisc.pr)
```
  
We need to make our own PCA plt
```{r}
#Make sure to put PC1 and PC2 so that they will be run against each other in x and y axis instead of being run against any other component
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis, 
     xlab = "PC1", ylab = "PC2")

#CUrrently "diagnosis" is composed of 0s and 1s which count for white and black but if you make it worth "diagnosis+1", it will be composed of 1s and 2s so now it will be red and black ---- but I specified the column so it colours it red and black 
  #so this would be "col=diagnosis+1"
```

Now to make it PC1 and PC3 (instead of PC2)
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,3], col=diagnosis+2, 
     xlab = "PC1", ylab = "PC3")

```

###Variance captured in each PC

This info is in the $sdev component of our PCA result
```{r}
# Calculate variance of each component
pr.var <- wisc.pr$sdev^2
head(pr.var)
```

```{r}
# Variance explained by each principal component: pve
variance <-  wisc.pr$sdev^2
pve <- round((variance/sum(variance))*100, 1)


# Plot variance explained for each principal component
plot(pve, type="o")
```

```{r}
barplot(pve)
```

Make the plot a bit fancy
```{r}
# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2))

#paste0("PC",1:length(pve)) gives us no space between the PC and the number where paste("PC",1:length(pve)) would give a normal space between PC and number
```

###Hierarchial clustering of case data

For hierarchial clustering, we need a few things
1.) Distance matrix **dist()** function
2.) The **hclust()** function
3.) Use the **cutree()** function to find cluster membership vector

```{r}
# Scale the wisc.data data: data.scaled
data.scaled <- scale(wisc.data)
```

```{r}
round(apply(wisc.data, 2, sd),1)
```

```{r}
round(apply(data.scaled, 2, sd),1)
```

Looks like we need scalling!!

Calculate the (Euclidean) distances between all pairs of observations in the new scaled dataset and assign the result to data.dist.
```{r}
data.dist <- dist(data.scaled)
```

```{r}
wisc.hclust <- hclust(data.dist,method = "complete")
plot(wisc.hclust)
```
 (skipped some methods in this lab period - go back through them for practice)
 
###Combining Methods

For clustering we need?
1.) Distance matrix
2.) Clustering function
3.) Cutree

This was our PCA result of PC1 vs PC2
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis, 
     xlab = "PC1", ylab = "PC2")
```

```{r}
pc.dist <- dist(wisc.pr$x[, 1:2])
pc.hclust <- hclust(pc.dist, method="ward.D2")
plot(pc.hclust)
abline(h=55, col="red")
```

```{r}
grps3 <- cutree(pc.hclust, k=3)
grps3
table(grps3)
```

```{r}
#This cross tabulates so we know how many of each of the diagnosis are in each of the groups. 0=benign, 1=malignant
table(grps3, diagnosis)
```

```{r}
#now colouring it in three groups that were defined in the previous step
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=grps3, 
     xlab = "PC1", ylab = "PC2")
```

###Trying to make the 3D Model

```{r}
library(rgl)
plot3d(wisc.pr$x[,1:3], xlab="PC 1", ylab="PC 2", zlab="PC 3", cex=1.5, size=1, type="s", col=diagnosis+1)
rglwidget(width = 400, height = 400)
```



##Prediction section
```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc

plot(wisc.pr$x[,1:2], col=diagnosis+1)
points(npc[,1], npc[,2], col="orange", pch=16, cex=2)
#find out which one may have cancer and which one may not have cancer based on the previous trends to make prediction
```























