---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Asthma SNPs in MXL population

Read file from Ensemle

http://uswest.ensembl.org/Homo_sapiens/Variation/Sample?db=core;r=17:39894595-39895595;v=rs8067378;vdb=variation;vf=362656008#373531_tablePanel

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
```

How many G|G genotypes are in this set?
```{r}
table(mxl$Genotype..forward.strand.)
sum(mxl$Genotype..forward.strand.=="G|G")
```

What proportion of the Mexican Ancestry in Los Angeles sample population (MXL) are homozygous for the asthma associated SNP (G|G)?
```{r}
table(mxl$Genotype..forward.strand.)/nrow(mxl) *100
```

##RNA-Seq

Q.Does the first sequence have good quality?

```{r}
#install.packages("seqinr")
#install.packages("gtools")
q <- "DDDDCDEDCDDDDBBDDDCC@"
library(seqinr)
library(gtools)

asc(s2c(q))-33
```

## Population scale analysis

Does the genotype for rs8067378 influence expression of the imprtant asthema linkd genes we found?
```{r}
geno <-read.table("rs8067378_ENSG00000172057.6.txt")
head(geno)
```
```{r}
table(geno$geno)
```
```{r}
inds.gg <- geno$geno =="G/G"
summary( geno$exp[inds.gg] )
```

```{r}
inds.aa <- geno$geno =="A/A"
summary( geno$exp[inds.aa] )
```
```{r}
inds.ag <- geno$geno =="A/G"
summary( geno$exp[inds.ag] )
```

Some nice figure that shows the distribution of these espression values?
```{r}
boxplot(exp ~ geno,geno, xlab="geno", ylab="exp")
```
```

