---
title: "Class 6"
author: "Reshma Pyala"
date: "1/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Section 1: Reading files again

Here we are going to try using the **read.table()** function and friends to read some example flat files.

First, let's try adding a simple plot

```{r}
plot(1:10)
```

*Back to file reading...*

```{r}
file1 <- ("https://bioboot.github.io/bimm143_W19/class-material/test1.txt")
read.table("https://bioboot.github.io/bimm143_W19/class-material/test1.txt", header = TRUE, sep = ",")
data1 <- read.csv(file1)

```


```{r}
file2 <- ("https://bioboot.github.io/bimm143_W19/class-material/test2.txt")
read.table("https://bioboot.github.io/bimm143_W19/class-material/test2.txt", header = TRUE, sep = "$")
data2 <- read.csv(file2)

```


```{r}
file3 <- ("test3.txt")
read.table(file3)
data3 <- read.csv(file3)
```


##Section 2: R Functions

My first silly function
```{r}
add <- function(x, y=1) {
# Sum the input x and y 
  x+y
}
```

Let's try ussing this function
```{r}
add(1)
```

```{r}
add(1, 100)
```

```{r}
add( c(1, 2, 3) )
```

```{r}
add( c(1, 2, 3), 4 )
```

```{r}
#add(1, 2, 2)
```

```{r}
#add(x=1, y=“b”)
```

##My 2nd function
```{r}
rescale <- function(x) {
   rng <-range(x)
   (x - rng[1]) / (rng[2] - rng[1])
}

```

Test on a small example where we know what the answer should be

```{r}
rescale(1:10)
```

```{r}
# How would you get your function to work here...
rescale( c(1,2,NA,3,10) )
```

```{r}

# What should your function do here?
rescale( c(1,10,"string") )
```

```{r}
rescale2 <- function(x) {
   rng <- range(x, na.rm = TRUE)
   (x - rng[1]) / (rng[2] - rng[1])
}

x <- ( c(1,2,NA,3,10) )
rng <- range(x)
#rng
```

```{r}
rescale2( c(1,2,NA,3,10) )
```

```{r}
rescale <- function(x, na.rm=TRUE, plot=FALSE) {
    rng <-range(x, na.rm=na.rm)
    print("Hello")
   answer <- (x - rng[1]) / (rng[2] - rng[1])
   print("is it me you are looking for?")
   if(plot) {
      plot(answer, typ="b", lwd=4)
   }
   print("I can see it in ...")
   return(answer)
}
```

```{r}
rescale3(c(1:6, NA, 10), plot=TRUE)
```

```{r}
rescale3(c(1:6, NA, 10), plot=FALSE)
```














##Working with the bio3d package

To install this package, I used the commang **install.packages("bio3d")**

```{r}
library(bio3d)
```

```{r}
#Read a PDB file from the database
s1 <- read.pdb("4AKE")  # kinase with drug
s1
```

Try out the provided code...
```{r}
s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```


















