---
title: "Class 07"
author: "Reshma Pyala"
date: "1/29/2019"
output:github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Functions revist
```{r}
source("http://tinyurl.com/rescale-R")
```

Let's try the rescale() function out
```{r}
rescale(c(1, 5, 10))
```
Looking to see what the error message is for rescale()
```{r}
rescale(c(1:5, "string"))
```


Let's try **rescale2()** with the **stop()** function catch for non-numeric input
```{r}
rescale2(c(1:5, "string"))
```
Using code to find NA in the data given.
```{r}
x <- c(3, 7, NA, 4, 8, NA)
is.na(x)
which(is.na(x))
```

```{r}
#Let's define an example x and y
x <- c(1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```

```{r}
is.na(x)
sum(is.na(x))
```

```{r}
is.na(x)
is.na(y)
is.na(x) & is.na(y)

#Putting together!
#sum(is.na(x) & is.na(y) - this tells you how many positions is it true for both x and y to have a NA

sum(is.na(x) & is.na(y))
```

Take my working snippet and make our first function...
```{r}
# No further simplification necessary
both_na <- function(x, y) {
  sum( is.na(x) & is.na(y) )
}
```

```{r}
both_na(x, y)
```

```{r}
#Trying with more variables to see if it works
x <-  c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

y3 <- c(1, NA, NA, NA, NA)
```

```{r}
both_na(x, y1)
```

```{r}
# What will this return?
both_na(x, y2)

#Now trying it with y3, which has an extra value
both_na(x, y3)
```

The reason that it is still reading a code above even when the second code is longer is because R recycles the script. So the first NA in x is recycles to the end of the code, which then makes it 4 characters long like y2. Then the NA and NA match up in position 4 which then gives us the reading of 3 matching NAs.
This is the same with y3. It reads as 4 NAs matching because the first NA of x is recycles to match with the fourth and fifth characters from y3. 


```{r}
#Adding a function to both_na2 to give us more specific error to let us know what to fix.
both_na2 <- function(x, y) {
  if(length(x) != length(y)) {
   stop("Input x and y should be the same length")
}
  sum( is.na(x) & is.na(y) )
}
```

```{r}
both_na2(x,y2)
```

```{r}
#New code added as both_na3 to give us a more specific response
both_na3(x, y1)
```

